---
title: "midterm project"
author: "Zhe Yu, Rong Li"
date: March 5
output: 
  revealjs::revealjs_presentation:
    theme: serif
    transition: slide
    self_contained: false
    reveal_plugins: ["menu"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(usmap, dplyr, tidyverse, drat, hurricaneexposuredata, ggplot2, maps, tmap, leaflet, sf, sp, lubridate, readr, kableExtra, magrittr, gridExtra, knitr)
```

# Introduction

In this assignment, we use additional county level data from FEMA to produce maps with more features. The purpose of this assignment is to show our ability of using R to produce documents, presentations, and shiny apps.  


# Data Collection

## Data Description

The Public Assistance Funded Projects Details dataset contains obligated (financial obligation to grantee) Public Assistance projects, lists public assistance recipients designated as applicants in the data, and a list of every funded, individual project, called project worksheets. Open projects still under pre-obligation processing are not represented. This is raw, unedited data from FEMA's National Emergency Management Information System (NEMIS). 

## Load Data

- We download raw data [PublicAssistanceFundedProjectsDetails.csv]( https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1) from the website of FEMA and load into R.  
Here are the first several rows and columns of the data:  
```{r, message = FALSE, echo = FALSE}
PAFPD <- read.csv("PublicAssistanceFundedProjectsDetails.csv",header = T)
as.data.frame(PAFPD[1:5, 1:7])
```
There are 79192 rows and 22 columns.  

- Obtain Hurricane Data from 2009 to 2018

```{r, echo = FALSE}
PAFPD$beginyear <- year(PAFPD$declarationDate)
PAFPD_0918 <- filter(PAFPD,beginyear > 2008 & beginyear < 2019 & incidentType == "Hurricane")
PAFPD_0918$stateNumberCode <- as.numeric(PAFPD_0918$stateNumberCode)
PAFPD_0918$countyCode <- as.numeric(PAFPD_0918$countyCode)
PAFPD_0918$fips <- 1000*PAFPD_0918$stateNumberCode+PAFPD_0918$countyCode
PAFPD_0918 <- filter(PAFPD_0918,projectAmount>=0)
as.data.frame(PAFPD_0918[1:5, 1:7])
```

# EDA

## Total Project Amount for Each State

```{r, echo = FALSE}
##barplot of the top 5 projectAmount in states
statesum <- PAFPD_0918 %>% group_by(state) %>% summarize(sum = sum(projectAmount))
kable(statesum[rev(order(statesum$sum))[1:10],])
p1 <- ggplot(data = statesum[rev(order(statesum$sum))[1:10],], mapping = aes(x = state, y = sum, fill = state)) + 
  geom_bar(stat = "identity") + 
  guides(fill = F) + 
  ylab("sum of projectAmount") + 
  ggtitle("Barplot of the top 10 projectAmount in states") + 
  theme(plot.title = element_text(hjust = 0.5))

##barplot of the other 22 projectAmount in states
p2 <- ggplot(data = statesum[rev(order(statesum$sum))[11:27],], mapping = aes(x = state, y = sum, fill = state)) + 
  geom_bar(stat = "identity") + 
  guides(fill = F) + 
  coord_flip() + 
  ylab("sum of projectAmount") + 
  ggtitle("Barplot of projectAmount for other states") + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1,p2,nrow=1)
```

## Total Project Amount for Each Year

```{r, echo = FALSE}
##barplot of the top 2 projectAmount in years
yearsum <- PAFPD_0918 %>% group_by(beginyear) %>% summarize(sum = sum(projectAmount))
kable(yearsum[rev(order(yearsum$sum)),])
p1 <- ggplot(data = yearsum[rev(order(yearsum$sum))[1:2],], mapping = aes(x = factor(beginyear), y = sum ,fill = factor(beginyear))) + 
  geom_bar(stat = "identity") + 
  guides(fill = F) + 
  xlab("year") + 
  ylab("sum of projectAmount") + 
  ggtitle("Barplot of the top 2 projectAmount in years") + 
  theme(plot.title = element_text(hjust = 0.5))
##barplot of the other projectAmount in years
p2 <- ggplot(data = yearsum[rev(order(yearsum$sum))[3:5],], mapping = aes(x = factor(beginyear), y = sum, fill = factor(beginyear))) + 
  geom_bar(stat = "identity") + 
  guides(fill = F) + 
  xlab("year") + 
  ylab("sum of projectAmount") + 
  ggtitle("Barplot of the No.3-No.5 for each year") + 
  theme(plot.title = element_text(hjust = 0.5))
p3 <- ggplot(data = yearsum[rev(order(yearsum$sum))[6:8],], mapping = aes(x = factor(beginyear), y = sum, fill = factor(beginyear))) + 
  geom_bar(stat = "identity") + 
  guides(fill = F) + 
  xlab("year") + 
  ylab("sum of projectAmount") + 
  ggtitle("Barplot of the other projectAmount for each year") + 
  theme(plot.title = element_text(hjust = 0.5))

grid.arrange(p1,p2,p3,nrow=1)
```

## Total Project Amount for Different Project Size

```{r, echo = FALSE}
##sum of projectAmount for different projectSize
sizesum <- PAFPD_0918 %>% group_by(projectSize) %>% summarize(sum = sum(projectAmount))
sizecount <- PAFPD_0918 %>% group_by(projectSize) %>% summarize(count = n(),.groups='drop')
kable(sizesum)
kable(sizecount)
##barplot of projectAmount for different projectSize
p1 <- ggplot(data = sizesum, mapping = aes(x = factor(projectSize), y = sum, fill = factor(projectSize))) + 
  geom_bar(stat = "identity") + 
  guides(fill = F) + 
  xlab("projectSize") + 
  ylab("sum of projectAmount") + 
  ggtitle("Barplot of projectAmount for different projectSize") + 
  theme(plot.title = element_text(hjust = 0.5))

p2 <- ggplot(data = sizecount, mapping = aes(x = factor(projectSize), y = count, fill = factor(projectSize))) + 
  geom_bar(stat = "identity") + 
  guides(fill = F) + 
  xlab("projectSize") + 
  ylab("count of projectAmount") + 
  ggtitle("Barplot of count for different projectSize") + 
  theme(plot.title = element_text(hjust = 0.5))

print(p1)
print(p2)

```

## Total Project Amount for Different Damage Categories

```{r, echo = FALSE}
##sum of projectAmount for different damageCategory
dccsum <- PAFPD_0918 %>% group_by(dcc) %>% summarize(sum = sum(projectAmount))
kable(dccsum[rev(order(dccsum$sum)),])
##Barplot of projectAmount for different damageCategory
ggplot(data = dccsum, mapping = aes(x = dcc, y = sum, fill = dcc)) + 
  geom_bar(stat = "identity") + 
  guides(fill=guide_legend(title=NULL)) + 
  scale_fill_discrete(labels=c("A-Debris Utilities", "B-Protective Measures", "C-Roads and Bridges", "D-Water Control Facilities", "E-Public Buildings", "F-Public Utilities", "G-Recreational or Other", "Z-State Management")) + 
  xlab("damageCatrgory") + 
  ylab("sum of projectAmount") + 
  ggtitle("Barplot of projectAmount for different damageCategory") + 
  theme(plot.title = element_text(hjust = 0.5))
```

# Mapping with 'Leaflet'

## Combine Fips Code with Maps Data

## Maps of States

We draw the total cost of every state in map.  

```{r echo=FALSE, message=FALSE}
#combine the fips with maps data
##county
data(county.fips)
mapcounty=st_as_sf(map('county',plot=F,fill=T))
colnames(county.fips)[2]=colnames(mapcounty)[1]
mapcounty=left_join(mapcounty,county.fips,'ID')

##state
data(state.fips)
mapstate=st_as_sf(map('state',plot = F,fill = T))
colnames(state.fips)[6]=colnames(mapstate)[1]
mapstate=left_join(mapstate,state.fips,'ID')
mapstate[20,2:6]<-c(25,22,1,1,"MA")
mapstate[21,2:6]<-c(26,23,2,3,"MI")
mapstate[31,2:6]<-c(36,33,1,2,"NY")
mapstate[32,2:6]<-c(37,34,3,5,"NC")
mapstate[45,2:6]<-c(51,49,3,5,"VA")
mapstate[46,2:6]<-c(53,50,4,9,"WA")
colnames(mapstate)[2] <- "stateNumberCode"
mapstate$stateNumberCode<- as.numeric(mapstate$stateNumberCode)

#combine data
#by state
hurricane_group<- group_by(PAFPD_0918, stateNumberCode)
h_GroupByfips<- summarise(hurricane_group,sumcost = sum(projectAmount))
h_GroupByfips<- h_GroupByfips[order(h_GroupByfips$sumcost,decreasing=F),]
hcount_s <- right_join(mapstate,h_GroupByfips,"stateNumberCode")

#by county
hurricane_group <- filter(PAFPD_0918,countyCode!=0)
hurricane_group<- group_by(hurricane_group, fips)
h_GroupByfips<- summarise(hurricane_group,sumcost = sum(projectAmount))
h_GroupByfips<- h_GroupByfips[order(h_GroupByfips$sumcost,decreasing=F),]
hcount_c <- right_join(mapcounty,h_GroupByfips,"fips")
hcount_c <- filter(hcount_c,is.na(ID)==F)
hcount_c2 <- filter(hcount_c,sumcost<200000000)
hcount_c3 <- filter(hcount_c,sumcost>200000000)
```

```{r, echo = FALSE}
#leaflet
#state
pal <- colorNumeric("YlOrRd",domain=hcount_s$sumcost)
pop <- paste("State:",hcount_s$ID,"<br/>",
             "estimated total cost:",hcount_s$sumcost,"$","<br/>")
leaflet(hcount_s)%>%
  addTiles()%>%
  addPolygons(data = hcount_s,fillColor=~pal(hcount_s$sumcost), 
              stroke = T,popup=pop,fillOpacity = 0.5,weight = 0.5,
              color = "gray")%>%
  addLegend("bottomright",values = ~sumcost,pal = pal,
            title = "estimated total cost")
```
  
As shown in the picture, hurricanes mainly influence the east coast. New York has the largest estimated total cost in 2009-2018.  

## Maps of Counties

If we put the top 10 counties with the largest estimated total cost in the same graph with other counties, the differences among counties will be hard to find.  
So we draw the top 10 counties alone and it appears that they are still hard to find in map because of their tiny area.  
As a result, we simply list them as follows:  

```{r, echo = FALSE}
rank10<-data.frame(county=1:10,funding=1:10)
rank10$county <- hcount_c3$ID
rank10$funding <- hcount_c3$sumcost
kable(rank10[order(rank10$funding,decreasing=T),])
```

And the rest counties are drawn in this map:  

```{r, echo = FALSE}
pal2 <- colorNumeric("YlOrRd",domain=hcount_c3$sumcost)
pop2 <- paste("County:",hcount_c3$ID,"<br/>",
              "estimated total cost:",hcount_c3$sumcost,"$","<br/>")
pal3 <- colorNumeric("PuRd",domain=hcount_c2$sumcost)
pop3 <- paste("County:",hcount_c2$ID,"<br/>",
              "estimated total cost:",hcount_c2$sumcost,"$","<br/>")
leaflet(hcount_c2)%>%
  addTiles()%>%
  addPolygons(data = hcount_c2,fillColor=~pal3(hcount_c2$sumcost), 
              stroke = T,popup=pop2,fillOpacity = 0.7,weight = 0.5,
              color = "gray")%>%
  addLegend("bottomright",values = ~sumcost,pal = pal3,
            title = "estimated total cost")

```

# Shiny

[Click here to shiny app](https://yuzhe.shinyapps.io/MA615-midtermmap/) 